Comparative Analysis of the wav2vec 2.0 Feature Extractor

### 2023-03-09 main.tex
add text for first layer freq resp

\subsection{Frequency Response of Learned Filters}
Finally, we analyze the learned filters of the neural \fes.
First, the frequency response of the filters in the first layer which operates on the raw waveform is plotted.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a) as a reference.
\refFig{fig:first_layer}\,(b) shows the frequency response of the learned \gls{SC} filters.
In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
This confirms, that these filters are also preferred by more advanced neural back-ends like self-attention-based like \conformer which are superior in the exploitation of context information compared to simpler neural back-ends e.g. in \cite{tuske2014raw}.
The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.
However, the stopband attenuation is clearly weaker.

The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}\,(c) does not entail this degree of resemblance.
Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
% for SCF, first 9 filters have peak at 0Hz, for wav2vec 75 filters

>>>>

\subsection{Frequency Response of Learned Filters}
Finally, we analyze the learned filters of the neural \fes.
First, the frequency response of the filters in the first layer which operates on the raw waveform is plotted.
As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria, respectively.
The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a) as a reference.
\refFig{fig:first_layer}\,(b) shows the frequency response of the learned \gls{SC} filters.
In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
This confirms, that these filters are also preferred by more advanced neural back-ends like self-attention-based like \conformer which are superior in the exploitation of context information compared to simpler neural back-ends e.g. in \cite{tuske2014raw}.
The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.
However, the stopband attenuation is clearly weaker and there is also a number of filters with no clear passband and a rather uncharacteristic frequency response.
This is particularly visible in \refFig{fig:first_layer}\,(e).

The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}\,(c) does not entail this degree of resemblance.
Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
Moreover, a number of filters exists with multiple passbands.
Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
